---
title: "607 week5 assign"
author: "Chris Ayre"
date: "2/28/2019"
output: html_document
---

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(janitor)
```


#### (1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

```{r}
numbersense <- read_csv("~/Google Drive/CUNY SPRING 19/COURSES/data 607/ASSIGNMENTS/numbersense.csv")
numbersense
```

#### (2) Read the information from your.CSV file into R,and use tidyr and dplyr as needed to tidy and transform your data.

Condensing the destinations(observations) to rows and  removing empty data.
```{r}
numbersense %>% remove_empty("rows")

numbersense_gather <- numbersense %>% gather(Destination, Flights, -X1, -X2, na.rm =TRUE)
numbersense_gather
```


rename columns for clarity
```{r}
names(numbersense_gather) <- c("AIRLINE", "Flight_Status", "Destination", "Flights")

```

spread function to focus on the target data 
```{r}
numbersense_spread <- numbersense_gather %>%
  spread(Flight_Status, Flights)
```

rename columns for clarity
```{r}
names(numbersense_spread) <- c("AIRLINE", "Destination", "Delayed", "On_Time")

```


##### (3) Perform analysis to compare the arrival delays for the two airlines.

Display the average number of delays for both airlines

```{r}
numbersense_spread %>%
  group_by(AIRLINE) %>%
  summarise(aver_delay_per_airline = mean(Delayed))
```
__AM WEST experiences 57% more delays than ALASKA__


Display the chance that a flight will be delayed based on the chosen Airline and Destination

```{r}
numbersense_spread %>%
  group_by(AIRLINE, Destination) %>%
  select(Delayed, On_Time) %>%
  mutate('(%) Chance_Delay' = (Delayed/(Delayed + On_Time)*100))
```
__With this information potential passengers will be see the probability of their flight being delayed before they book it.__


Display the two destinations that experience the lowest number of delays per airline.

```{r}
numbersense_spread %>%
  group_by(AIRLINE,Destination) %>%
  summarise(min_delay_per_dest = min(Delayed)) %>%
  arrange(Destination) %>% slice(1:2)
```
__Both Airlines experience their lowest number of delays in LA and Phoenix However, ALASKA is 417% more likely to not experience a delay in Phoenix, while AM WEST is 255% more likely to not experience a delay in LA__

